---
title: "Exercise 6"
subtitle: "for Advanced Methods for Regression and Classification"
author: "Eszter Katalin Bognar"
date: "18.11.2020"
output: pdf_document
---

To prevent cutting the pdf
```{r, echo = TRUE}
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)
```

# Olitos dataset

Load and view the data
```{r, echo=TRUE}
data(olitos,package="rrcov")
str(olitos)
```

Performing principal component analysis for the 25 x-variables
```{r, echo=TRUE}
olitos.pca <- prcomp(olitos[,c(1:25)], center = TRUE,scale. = TRUE)
summary(olitos.pca)
str(olitos.pca)
```

Show a plot of the first 2 PCA scores, with color information for the 4 groups
```{r, echo=TRUE,fig.width=6,fig.height=4}
olitosClasses <- factor(olitos$grp)
plot(olitos.pca$x[,1:2], col = olitosClasses)
```

Based on the plot, we can see that using two components it hard to find a clear decision boundary between the 4 classes.

Train/test split
```{r, echo=TRUE}
set.seed(123)
smp_size <- floor(2/3* nrow(olitos))
train_ind <- sample(seq_len(nrow(olitos)), size = smp_size)
```

Apply LS
```{r, echo=TRUE}
alias(as.numeric(grp)~.,data=olitos)
model.ls <- lm(as.numeric(grp)~.,data=olitos,subset=train_ind)
summary(model.ls)
```

By looking at the model summary we can see, that only the intercept and a few variables are contributing significantly to the model, X2 has the highest significance and X13 and X23 some lower significance.

Make predictions, Binary response matrix (confusion matrix)
```{r, echo=TRUE}
library(caret)
pred <- predict(model.ls,olitos[train_ind,])
predfact = cut(pred, 4, labels=c('1', '2', '3','4'))
confusionMatrix(olitos$grp[train_ind],predfact)
```

Evaluation on the test set
```{r, echo=TRUE}
library(caret)
pred <- predict(model.ls,olitos[-train_ind,])
predfact = cut(pred, 4, labels=c('1', '2', '3','4'))
confusionMatrix(olitos$grp[-train_ind],predfact)
```
The model cannot distinguish between the classes, specially on the test set. There are misclassifications occurring mostly for groups 2,3,4, group 1 is identified better. This corresponds with the first plot, where the "green" class has a clearer decision boundary.

# Bank + Marketing dataset

Load the data
```{r, echo=TRUE}
# bank data set:
d <- read.csv2("bank.csv")
table(d$y)
```

Select randomly a training set with 3000 observations
```{r, echo=TRUE}
n <- nrow(d)
set.seed(123)
train <- sample(1:n,3000)
test <- c(1:n)[-train]
```

use lm() to predict the group label
```{r, echo=TRUE}
model.lm <- lm(as.numeric(y)~.,data=d,subset=train)
pred <- predict(model.lm,d[test,])
library(caret)
predfact = cut(pred, 2, labels=c('no', 'yes'))
confusionMatrix(d$y[test],predfact)
```

Apply undersampling
```{r, echo=TRUE}
table(d$y[train])
train.yes <- train[d$y[train]=="yes"]
train.no <- sample(train[d$y[train]=="no"],length(train.yes))
model.lm_undersampling <- lm(as.numeric(y)~.,data=d,subset=c(train.yes,train.no))
pred_undersampling <- predict(model.lm_undersampling,d[test,])
```

Evaluate the results
```{r, echo=TRUE}
library(caret)
predfact = cut(pred_undersampling, 2, labels=c('no', 'yes'))
confusionMatrix(d$y[test],predfact)
```

The accuracy increased and the misclassification error decreased.

Apply oversampling
```{r, echo=TRUE}
table(d$y[train])
train.no <- train[d$y[train]=="no"]
train.yes <- sample(train[d$y[train]=="yes"],length(train.no),
                    replace=TRUE)
model.lm_oversampling <- lm(as.numeric(y)~.,data=d,subset=c(train.yes,train.no))
pred_oversampling <- predict(model.lm_oversampling,d[test,])
```

Evaluate the results
```{r, echo=TRUE}
library(caret)
predfact = cut(pred_oversampling, 2, labels=c('no', 'yes'))
confusionMatrix(d$y[test],predfact)
```

With oversampling, the results are worse.