---
title: "StatComp_11931695"
subtitle: "for Exercises"
author: "Eszter Katalin Bognar"
date: "19-05-2020"
output: pdf_document
---

# Introduction to Data Visualization with ggplot2
## mapping onto a color gradient or onto a continuous size scale or shape (only for categorial variables)

```{r, echo=TRUE,fig.width=6,fig.height=4}
library(ggplot2)
ggplot(mtcars, aes(x = wt, y = mpg, color = disp)) + geom_point()
ggplot(mtcars, aes(x = wt, y = mpg, size = disp)) + geom_point()
```
## Adding geometries
Common geom layer functions:
-geom_point() adds points (as in a scatter plot)
-geom_smooth() adds a smooth trend curve
We can also make change to individual geoms, such as making the points 40% opaque
```{r, echo=TRUE,fig.width=6,fig.height=4}
ggplot(diamonds, aes(x = carat, y = price, color = clarity)) +
  geom_point(alpha=0.4) +
  geom_smooth()
```
## Make changes to individual geoms and saving plots as variables
```{r, echo=TRUE,fig.width=6,fig.height=4}
plt_price_vs_carat <- ggplot(diamonds, aes(carat, price))

# Edit this to map color to clarity,
# Assign the updated plot to a new object
plt_price_vs_carat_by_clarity <- plt_price_vs_carat + geom_point(alpha = 0.2) + geom_smooth(aes(col = clarity), se = F)

# See the plot
plt_price_vs_carat_by_clarity
```
## Visible aesthetics -  color, fill, shape and size

```{r, echo=TRUE,fig.width=6,fig.height=4}
# Map cyl to fill and 
ggplot(mtcars, aes(x = wt, y = mpg, fill=gear, color = factor(cyl))) +
  geom_point(shape = 21, size = 4)

```
## Using geom_text
```{r, echo=TRUE,fig.width=6,fig.height=4}
plt_mpg_vs_wt <- ggplot(mtcars, aes(wt, mpg))

# Use text layer and map fcyl to label
plt_mpg_vs_wt +
  geom_text(aes(label = cyl))
```
## Updating aesthetic labels
```{r, echo=TRUE,fig.width=6,fig.height=4}
ggplot(mtcars, aes(cyl, fill = as.factor(am))) +
  geom_bar() +
  # Set the axis labels
  labs(x = "Number of Cylinders", y = "Count")

```
## Using custom color scale
```{r, echo=TRUE,fig.width=6,fig.height=4}
palette <- c(automatic = "#377EB8", manual = "#E41A1C")

# Set the position
ggplot(mtcars, aes(factor(cyl), fill = factor(am))) +
  geom_bar(position = "dodge") +
  labs(x = "Number of Cylinders", y = "Count")
  scale_fill_manual("Transmission", values = palette)
```
## Setting a dummy aesthetic
```{r, echo=TRUE,fig.width=6,fig.height=4}
ggplot(mtcars, aes(mpg, 0)) +
  geom_jitter() +
  # Set the y-axis limits
  ylim(limits = c(-2,2))

```
## Scatter plots

```{r, echo=TRUE,fig.width=6,fig.height=4}
# Plot price vs. carat, colored by clarity
plt_price_vs_carat_by_clarity <- ggplot(diamonds, aes(carat, price, color = clarity))

# Set transparency to 0.5
plt_price_vs_carat_by_clarity + geom_point(alpha = 0.5, shape = 16)

```
## Aligned values
```{r, echo=TRUE,fig.width=6,fig.height=4}
# Plot base
plt_mpg_vs_fcyl_by_fam <- ggplot(mtcars, aes(factor(cyl), mpg, color = factor(am)))

# Default points are shown for comparison
plt_mpg_vs_fcyl_by_fam + geom_point()

# Now jitter and dodge the point positions
plt_mpg_vs_fcyl_by_fam + geom_point(position = position_jitterdodge(jitter.width = 0.3, dodge.width = 0.3))

```
## Applying jittering and alpha blending as solution to the overplotting
```{r, echo=TRUE,fig.width=6,fig.height=4}
library(car)
ggplot(Vocab, aes(education, vocabulary)) +
  # Set the shape to 1
  geom_jitter(alpha = 0.2, shape=1)
```
## Histograms

```{r, echo=TRUE,fig.width=6,fig.height=4}
datacamp_light_blue <- "#51A8C9"

ggplot(mtcars, aes(mpg, ..density..)) +
  # Set the fill color to datacamp_light_blue
  geom_histogram(binwidth = 1, fill=datacamp_light_blue)

```
## Putting bars on top of eachother
```{r, echo=TRUE,fig.width=6,fig.height=4}
ggplot(mtcars, aes(mpg, fill = factor(am))) +
  # Change the position to identity, with transparency 0.4
  geom_histogram(binwidth = 1, position = "identity", alpha=0.4)
```
## Bar plots

```{r, echo=TRUE,fig.width=6,fig.height=4}
ggplot(mtcars, aes(factor(cyl), fill = factor(am))) +
  # Change the position to "dodge"
  geom_bar(position = "dodge")

```
## Overlapping bar plots
```{r, echo=TRUE,fig.width=6,fig.height=4}
ggplot(mtcars, aes(cyl, fill = factor(am))) +
  # Set the transparency to 0.6
  geom_bar(position = position_dodge(width = 0.2),alpha=0.6)

```
## Line plots

```{r, echo=TRUE,fig.width=6,fig.height=4}
# Change the y-axis to the proportion of the population that is unemployed
ggplot(economics, aes(date,unemploy/pop)) +
  geom_line()
```
## Themes layer

```{r, echo=TRUE,fig.width=6,fig.height=4}
# Position the legend inside the plot at (0.6, 0.1)
p=ggplot(ChickWeight, 
    aes(x = Time, y = weight, col = Diet)) + 
  geom_line(
    aes(group = Chick), alpha = 0.3) +
  theme(legend.position = "left")
```
## Modifying theme elements
```{r, echo=TRUE,fig.width=6,fig.height=4}
p +
  theme(
    rect = element_rect(fill = "grey92"),
    legend.key = element_rect(color = NA),
    axis.ticks = element_blank(),
    panel.grid = element_blank(),
    panel.grid.major.y = element_line(
      color = "white",
      size = 0.5,
      linetype = "dotted"
    ),
    # Set the axis text color to grey25
    axis.text = element_text(color = "grey25"),
    # Set the plot title font face to italic and font size to 16
   plot.title=element_text(size=16, face="italic")
  )
```
## Modifying whitespace
```{r, echo=TRUE,fig.width=6,fig.height=4}

p +
  theme(
    # Set the axis tick length to 2 lines
    axis.ticks.length = unit(2,"lines")
  )
```
```{r, echo=TRUE,fig.width=6,fig.height=4}

p +
  theme(
    # Set the plot margin to (10, 30, 50, 70) millimeters
    plot.margin = margin(10, 30, 50, 70, "mm")
  )
```
## Built-in themes

```{r, echo=TRUE,fig.width=6,fig.height=4}
# Add a void theme
p +
  theme_void()
# Add a classic theme
p +
  theme_classic()
# Add a bw theme
p +
  theme_bw()

```
# Intermediate Data Visualitatiton with ggplot2 
## Statistics

```{r, echo=TRUE,fig.width=6,fig.height=4}
ggplot(mtcars, 
    aes(x = wt, y = mpg, color = factor(cyl))) +
  # Add a point layer
  geom_point() +
  # Add a smooth lin reg stat, no ribbon
  stat_smooth(method="lm",se=FALSE)

```
## Amend the plot
```{r, echo=TRUE,fig.width=6,fig.height=4}
# Amend the plot
ggplot(mtcars, aes(x = wt, y = mpg, color = factor(cyl))) +
  geom_point() +
  # Map color to dummy variable "All"
  stat_smooth(se = FALSE,aes(color="All")) +
  stat_smooth(method = "lm", se = FALSE)
```
## Quantile
```{r, echo=TRUE,fig.width=6,fig.height=4}
ggplot(Vocab, aes(x = education, y = vocabulary)) +
  geom_jitter(alpha = 0.25) +
  stat_quantile(quantiles = c(0.05, 0.5, 0.95))
```
## Sum
```{r, echo=TRUE,fig.width=6,fig.height=4}
ggplot(Vocab, aes(x = education, y = vocabulary, group = education)) +
  stat_sum(aes(size = ..prop..))
```
## Using stats outside geoms
```{r, echo=TRUE,fig.width=6,fig.height=4}
posn_j <- position_jitter(width = 0.2)
posn_d <- position_dodge(width = 0.1)
posn_jd <- position_jitterdodge(jitter.width = 0.2, dodge.width = 0.1)

# Create the plot base: wt vs. fcyl, colored by fam
p_wt_vs_fcyl_by_fam_jit <- ggplot(mtcars,aes(factor(cyl),wt,color=factor(am)))

# Add a point layer
p_wt_vs_fcyl_by_fam_jit +
  geom_point()
```
## Using the stat_summary function
```{r, echo=TRUE,fig.width=6,fig.height=4}
p_wt_vs_fcyl_by_fam_jit +
  # Add a summary stat of std deviation limits
  stat_summary(fun.data = mean_sdl,
  fun.args = list(mult = 1),
  position = posn_d)

```
## Coordinates layer

```{r, echo=TRUE,fig.width=6,fig.height=4}
ggplot(mtcars, aes(x = wt, y = hp, color = factor(am))) +
  geom_point() +
  geom_smooth() +
  # Add Cartesian coordinates with x limits from 3 to 6
  coord_cartesian(xlim = c(3,6))
```
## Setting aspect ratio
```{r, echo=TRUE,fig.width=6,fig.height=4}
ggplot(iris, aes(x = Sepal.Length, y = Sepal.Width, color = Species)) +
  geom_jitter() +
  geom_smooth(method = "lm", se = FALSE) +
  # Fix the coordinate ratio
  coord_equal()
```
## Expand and clip
```{r, echo=TRUE,fig.width=6,fig.height=4}
ggplot(mtcars, aes(wt, mpg)) +
  geom_point(size = 2) +
  # Turn clipping off
  coord_cartesian(expand = 0, clip="off") +
  theme_classic() +
  # Remove axis lines
  theme(axis.line=element_blank())
```
## Log-transforming scales
```{r, echo=TRUE,fig.width=6,fig.height=4}
# Perform a log10 coordinate system transformation
ggplot(msleep, aes(bodywt, brainwt)) +
  geom_point() +
  coord_trans(x="log10",y="log10")
```
# Log-transforming scales
```{r, echo=TRUE,fig.width=6,fig.height=4}
# Plot with a scale_*_*() function:
ggplot(msleep, aes(bodywt, brainwt)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  # Add a log10 x scale
  scale_x_log10() +
  scale_y_log10() +
  ggtitle("Scale functions")
```
## Flipped axes
```{r, echo=TRUE,fig.width=6,fig.height=4}
ggplot(mtcars, aes(factor(cyl), fill = factor(am))) +
  # Set a dodge width of 0.5 for partially overlapping bars
  geom_bar(position = "dodge") +
  coord_flip()
```
## Pie chart - Polar coordinates

```{r, echo=TRUE,fig.width=6,fig.height=4}
ggplot(mtcars, aes(x = 1, fill = factor(cyl))) +
  geom_bar() +
  # Add a polar coordinate system
  coord_polar()
```
# Make some adjustment
```{r, echo=TRUE,fig.width=6,fig.height=4}
ggplot(mtcars, aes(x = 1, fill = factor(cyl))) +
  # Reduce the bar width to 0.1
  geom_bar(width=0.1) +
  coord_polar(theta = "y") +
  # Add a continuous x scale from 0.5 to 1.5
  scale_x_continuous(limits=c(0.5,1.5))
```
## The facets layer

```{r, echo=TRUE,fig.width=6,fig.height=4}
ggplot(mtcars, aes(wt, mpg)) + 
  geom_point() +
  # Facet rows by am and columns by cyl
   facet_grid(rows=vars(am),cols=vars(cyl))
```
## Facet formula formulation
```{r, echo=TRUE,fig.width=6,fig.height=4}
ggplot(mtcars, aes(wt, mpg)) + 
  geom_point() +
  # Facet rows by am using formula notation
  facet_grid(am ~ .) 

ggplot(mtcars, aes(wt, mpg)) + 
  geom_point() +
  # Facet columns by cyl using formula notation
  facet_grid(. ~ cyl)
  
ggplot(mtcars, aes(wt, mpg)) + 
  geom_point() +
  # Facet rows by am and columns by cyl using formula notation
  facet_grid(am ~ cyl)
```
## Labeling facets
```{r, echo=TRUE,fig.width=6,fig.height=4}
ggplot(mtcars, aes(wt, mpg)) +
  geom_point() +
  # Two variables
  facet_grid(cols = vars(vs, cyl), labeller = label_context)
```
## Labeling facets
```{r, echo=TRUE,fig.width=6,fig.height=4}
# Make factor, set proper labels explictly, and
# manually set the label order
mtcars$fam <- factor(mtcars$am,
                     levels = c(1, 0),
                     labels = c("manual", "automatic"))

# View again
ggplot(mtcars, aes(wt, mpg)) +
  geom_point() +
  facet_grid(cols = vars(fam))
```
## Scaling and spacing
```{r, echo=TRUE,fig.width=6,fig.height=4}
ggplot(mtcars, aes(x = mpg, y = factor(cyl), color = factor(am))) +
  geom_point() +
  # Free the y scales and space
  facet_grid(rows = vars(gear),space="free_y",scales="free_y")
```
##Facet wrapping for many levels
```{r, echo=TRUE,fig.width=6,fig.height=4}
ggplot(Vocab, aes(x = education, y = vocabulary)) +
  stat_smooth(method = "lm", se = FALSE) +
  # Create facets, wrapping by year, using vars()
  facet_wrap(vars(year))
```
## Margin plots
```{r, echo=TRUE,fig.width=6,fig.height=4}
ggplot(mtcars, aes(x = wt, y = mpg)) + 
  geom_point() +
  # Update the facets to only show margins on gear and fvs
  facet_grid(rows = vars(factor(vs), factor(am)), cols = vars(gear), margins = c("gear","fvs"))
```
## Best practices - Bar plots: dynamite plots

```{r, echo=TRUE,fig.width=6,fig.height=4}
# Define a dodge position object with width 0.9
posn_d <- position_dodge(width = 0.9)

# For each summary stat, update the position to posn_d
ggplot(mtcars, aes(x = factor(cyl), y = wt, color = factor(am), fill = factor(am))) +
  stat_summary(fun.y = mean, geom = "bar", position = posn_d, alpha = 0.5) +
  stat_summary(fun.data = mean_sdl, fun.args = list(mult = 1), width = 0.1, position = posn_d, geom = "errorbar")
```
## Heatmaps

```{r, echo=TRUE,fig.width=6,fig.height=4}
library(lattice)
ggplot(barley,aes(year,variety,fill=yield)) +
  # Add a tile geom
  geom_tile()
```
##Heatmap alternatives

```{r, echo=TRUE,fig.width=6,fig.height=4}
# Using barley, plot yield vs. year, colored and grouped by variety
ggplot(barley,aes(year,yield,group=variety,color=variety)) +
  # Add a line layer
  geom_line() +
  # Facet, wrapping by site, with 1 row
  facet_wrap( ~ site, nrow = 1)
```
</div></pre>